---
title: "Figure3C_Pathway_Gene_CircularPlot"
output: html_document
date: "2026-06-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Figure 3C Pathway–Gene Circular Visualization

This workflow documents the pathway–gene circular visualization shown in Figure 3C.

DAVID enrichment results were used to define enriched pathways, associated genes, and enrichment significance values. Pathway–gene relationships were visualized using the `circlize` package.

```{r libraries}
library(circlize)
library(dplyr)
library(readr)
```

```{r input_data}
# Input file containing DAVID-derived pathway-gene relationships.
#
# Required columns:
# Pathway, Gene, neg_log10_pvalue
#
# Example:
# Pathway,Gene,neg_log10_pvalue
# Cell adhesion,FN1,2.46
# Cell adhesion,ITGB1,2.46
# Wnt signaling pathway,WNT5A,1.49

pathway_gene <- read.csv("path_to_input/Figure3C_Pathway_Gene_Input.csv")
```

```{r preprocessing}
pathway_gene <- pathway_gene %>%
  dplyr::select(Pathway, Gene, neg_log10_pvalue) %>%
  dplyr::filter(!is.na(Pathway), !is.na(Gene)) %>%
  dplyr::distinct()

links <- pathway_gene %>%
  dplyr::select(Pathway, Gene) %>%
  dplyr::rename(
    from = Pathway,
    to = Gene
  )

pathways <- unique(links$from)
genes <- unique(links$to)
```

```{r colors}
# Sector colors distinguish pathway sectors from gene sectors.
grid_colors <- c(
  setNames(rep("#66C2A5", length(pathways)), pathways),
  setNames(rep("#FC8D62", length(genes)), genes)
)

# Enrichment values can be used for external annotation/color scaling
# during final figure assembly.
pathway_scores <- pathway_gene %>%
  dplyr::group_by(Pathway) %>%
  dplyr::summarise(
    neg_log10_pvalue = max(neg_log10_pvalue, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r circular_plot, fig.width=12, fig.height=12}
circos.clear()

circos.par(
  start.degree = 90,
  gap.degree = 2,
  track.margin = c(0.01, 0.01),
  points.overflow.warning = FALSE
)

chordDiagram(
  x = links,
  grid.col = grid_colors,
  transparency = 0.35,
  annotationTrack = "grid",
  preAllocateTracks = 1
)

circos.trackPlotRegion(
  track.index = 1,
  panel.fun = function(x, y) {

    sector.name <- get.cell.meta.data("sector.index")
    xlim <- get.cell.meta.data("xlim")
    ylim <- get.cell.meta.data("ylim")

    circos.text(
      x = mean(xlim),
      y = ylim[1],
      labels = sector.name,
      facing = "clockwise",
      niceFacing = TRUE,
      adj = c(0, 0.5),
      cex = 0.45
    )
  },
  bg.border = NA
)

circos.clear()
```

```{r save_output}
# Optional PDF export
pdf("Figure3C_Pathway_Gene_CircularPlot.pdf", width = 12, height = 12)

circos.clear()

circos.par(
  start.degree = 90,
  gap.degree = 2,
  track.margin = c(0.01, 0.01),
  points.overflow.warning = FALSE
)

chordDiagram(
  x = links,
  grid.col = grid_colors,
  transparency = 0.35,
  annotationTrack = "grid",
  preAllocateTracks = 1
)

circos.trackPlotRegion(
  track.index = 1,
  panel.fun = function(x, y) {

    sector.name <- get.cell.meta.data("sector.index")
    xlim <- get.cell.meta.data("xlim")
    ylim <- get.cell.meta.data("ylim")

    circos.text(
      x = mean(xlim),
      y = ylim[1],
      labels = sector.name,
      facing = "clockwise",
      niceFacing = TRUE,
      adj = c(0, 0.5),
      cex = 0.45
    )
  },
  bg.border = NA
)

dev.off()
circos.clear()
```

## Output

Figure 3C shows a circular pathway–gene visualization generated from DAVID enrichment output and pathway-associated gene lists using the `circlize` package.
